Differnces-in-Differences with Spatial Spillover

### 2020-12-04 latex/slides/Spillover_slides.tex
Econometrics Brownbag Slides

\begin{frame}{Spatial Spillovers}

    Researchers use difference-in-differences to estimate the average treatment effect on the treated: 
    \[
        \mathbb{E} \{ Y_{i1}(1) - Y_{i1}(0) \ \vert \ D_{i} = 1 \}
    \]

    Effects of treatments at states, counties, etc. can spill over boundaries.

@@

\begin{frame}{Spatial Spillovers}
    The difference-in-differences estimate is: 
    \[ 
        \hat{\mathbb{E}} \left[ Y_{i1} - Y_{i0} \mid D_i = 1 \right] - 
        \hat{\mathbb{E}} \left[ Y_{i1} - Y_{i0} \mid D_i = 0 \right]     
    \]

    Two problems in presence of spillover effects: 
    \begin{itemize}
        \item Nearby ``control'' units fail to estimate counterfactual trends because they are affected by treatment

        \item Treated units are also affected by nearby units and therefore combines ``direct'' effects with spillover effects
    \end{itemize}

@@

\begin{frame}{Research Question}
    What are the sources of bias when researchers do not account for sptial spillover of treatment effects?

    \begin{itemize}
        \item In what contexts are these biases particularly large?

        \item Do current simple adjustments work?
    \end{itemize}
\end{frame}

@@

        $h_i(\vec{D})$ = 1 if there is a treated unit within $x$-miles of unit $i$ and = 0 otherwise

        \item \textbf{Average treatment among k-nearest neighbors:}

@@

    \begin{itemize}
        \item In the complete absence of treatment (not just the absence of individual $i$'s treatment), the change in potential outcomes from period 0 to 1 would not depend on treatment status
    \end{itemize}
\end{frame}

@@

    \begin{align*}
        &\mathbb{E}_i \left[ Y_{i1} - Y_{i0} \mid D_i = 1 \right] - \mathbb{E}_i \left[ Y_{i1} - Y_{i0} \mid D_i = 0 \right] \\
        &\quad = 
        \textcolor{green}{\mathbb{E}_i \left[ Y_{i1}(1, 0) - Y_{i1}(0, 0) \mid D_i = 1 \right]} \\
        &\quad\quad + 
        \textcolor{red}{\mathbb{E}_i \left[ Y_{i1}(1, h_i(\vec{D})) - Y_{i1}(1, 0) \mid D_i = 1 \right]} \\ 
        &\quad\quad - 
        \textcolor{purple}{\mathbb{E}_i \left[ Y_{i1}(0, h_i(\vec{D})) - Y_{i1}(0, 0) \mid D_i = 0 \right]} \\
        &\quad \equiv \textcolor{green}{\tau_{\text{direct}}} + \textcolor{red}{\tau_{\text{spillover, treated}}} - \textcolor{purple}{\tau_{\text{spillover, control}}}
    \end{align*}

@@

    $\mathbb{E}_i \left[ Y_{i1}(1, h_i(\vec{D})) - Y_{i1}(1, 0) \mid D_i = 1 \right] = -0.5$ if county $i$ is within 40 miles of a treated county and $0$ otherwise.

@@

    $\mathbb{E}_i \left[ Y_{i1}(0, h_i(\vec{D})) - Y_{i1}(0, 0) \mid D_i = 0 \right] = 1$ if county $i$ is within 40 miles of a treated county and $0$ otherwise.

    \begin{itemize}
        \item Then, the bias term is \begin{align*}
            -0.5 * \text{\% of treated within 40 miles of a treated county} \\
            -1 * \text{\% of control within 40 miles of a treated county}
        \end{align*}
    \end{itemize}
\end{frame}

\begin{frame}{Partial Identification}
    Now, I turn to formalizing the above example following the literature on partial identification for treatment effects: \begin{citecolor}
        [\citet{Armstrong_Kolesár_2020}, \citet{Rambachan_Roth_2020}, and \citet{Manski_Pepper_2018}]
    \end{citecolor}

    In this framework, researchers need to specify bounds on spillovers in the context of their research question.
\end{frame}

@@

@@

@@

    They trim the sample using a logit regression to predict treatment using $X_{c,1940}$ and keep control units in the top 75\% of predicted probability.
\end{frame}

@@

                \input{../../tables/tva_replication_slides.tex}

                \bottomrule
            \end{tabular}

@@

>>>>

\begin{frame}{Spatial Spillovers}
    Researchers aim to estimate the average treatment effect on the treated: 
    \[
        \tau \equiv \mathbb{E} \left[ Y_{i1}(1) - Y_{i1}(0) \ \vert \ D_{i} = 1 \right]
    \]

    \vspace{5mm}
    Spillover effects are when effect of treatment extend over boundaries such as states, counties, etc.

@@

\begin{frame}{Bias from Spatial Spillovers}
    The canonical difference-in-differences estimate is: 
    \[ 
        \hat{\tau} = \underbrace{\hat{\mathbb{E}} \left[ Y_{i1} - Y_{i0} \mid D_i = 1 \right]}_{\text{Counterfactual Trend} \ + \ \tau} - 
        \underbrace{\hat{\mathbb{E}} \left[ Y_{i1} - Y_{i0} \mid D_i = 0 \right]}_{\text{Counterfactual Trend}}
    \]

    Two problems in presence of spillover effects: 
    \begin{itemize}
        \item \textbf{Spillover onto Control Units:}

        Nearby ``control'' units fail to estimate counterfactual trends because they are affected by treatment

        \item \textbf{Spillover onto other Treated Units:}

        Treated units are also affected by nearby units and therefore combines ``direct'' effects with spillover effects
    \end{itemize}

@@

\begin{frame}{Outline}

    \begin{itemize}
        \item[1--] Using a potential outcomes framework, I derive the bias in the presence of spillovers

        \item[2--] Using Monte Carlo simulations, show when biases are large and analyze effectiveness of current adjustments done by applied researchers

        \item[3--] Revisit \citet{Kline_Moretti_2014} analysis of the Tennessee Valley Authority
    \end{itemize} 
\end{frame}

@@

        $h_i(\vec{D})$ equals 1 if there is a treated unit within $x$-miles of unit $i$ and equals 0 otherwise

        \item \textbf{Average treatment among $k$-nearest neighbors:}

@@

    \vspace{5mm}
    In the complete absence of treatment (not just the absence of individual $i$'s treatment), the change in potential outcomes from period 0 to 1 would not depend on treatment status

@@

    \begin{align*}
        &\mathbb{E} \left[ Y_{i1} - Y_{i0} \mid D_i = 1 \right] - \mathbb{E} \left[ Y_{i1} - Y_{i0} \mid D_i = 0 \right] \\
        &\quad = 
        \textcolor{green}{\mathbb{E} \left[ Y_{i1}(1, 0) - Y_{i1}(0, 0) \mid D_i = 1 \right]} \\
        &\quad\quad + 
        \textcolor{red}{\mathbb{E} \left[ Y_{i1}(1, h_i(\vec{D})) - Y_{i1}(1, 0) \mid D_i = 1 \right]} \\ 
        &\quad\quad - 
        \textcolor{purple}{\mathbb{E} \left[ Y_{i1}(0, h_i(\vec{D})) - Y_{i1}(0, 0) \mid D_i = 0 \right]} \\
        &\quad \equiv \textcolor{green}{\tau_{\text{direct}}} + \textcolor{red}{\tau_{\text{spillover, treated}}} - \textcolor{purple}{\tau_{\text{spillover, control}}}
    \end{align*}

@@

    $Y_{i1}(1, h_i(\vec{D})) - Y_{i1}(1, 0) = -0.5$ if county $i$ shares a border with a treated county and $0$ otherwise.

@@

    $Y_{i1}(0, h_i(\vec{D})) - Y_{i1}(0, 0) \mid D_i = 1$ if county $i$ shares a border with a treated county and $0$ otherwise.

    Then, the bias term is 
    \begin{align*}
        &-0.5 * \text{\% of treated counties sharing a border with a treated county} \\
        &-1 * \text{\% of treated counties sharing a border with a treated county}
    \end{align*}
\end{frame}

\begin{frame}{Partial Identification}
    Researchers can us the above method to create bounds for the biases resulting from spillovers. The choices should follow from the empirical context at hand.

    To formalize this, I follow the literature on partial identification for treatment effects:

    \begin{citecolor}
        [\citet{Armstrong_Kolesár_2020}, \citet{Rambachan_Roth_2020}, and \citet{Manski_Pepper_2018}]
    \end{citecolor}
\end{frame}

@@

% ------------------------------------------------------------------------------
% \section{Estimation with Spillovers}
% ------------------------------------------------------------------------------

@@

% \begin{frame}{Including Spillovers in Regression}
%    How well does the modified specification do with misspecified spillovers?
% \end{frame}

@@

    They trim the sample using a logit regression to predict treatment using $X_{c,1940}$ and then keep control units in the top 75\% of predicted probability.
\end{frame}

@@

                Agricultural employment     & $-0.0514^{***}$& $-0.0678^{***}$& $-0.0310^{**}$ &    $-0.0112$   & $-0.0252^{***}$\\
                &   $(0.0087)$   &   $(0.0102)$   &   $(0.0123)$   &   $(0.0094)$   &   $(0.0084)$

                \onslide<2->{
                \\
                Manufacturing employment    & $0.0560^{***}$ &  $0.0461^{**}$ &    $-0.0104$   &    $-0.0128$   &  $-0.0248^{*}$ \\
                &   $(0.0187)$   &   $(0.0210)$   &   $(0.0205)$   &   $(0.0257)$   &   $(0.0147)$ 
                }
                \onslide<3->{
                \\
                Median family income        & $0.0191^{***}$ &  $0.0196^{**}$ &    $0.0056$    &    $-0.0062$   &    $-0.0011$   \\
                &   $(0.0067)$   &   $(0.0077)$   &   $(0.0079)$   &   $(0.0060)$   &   $(0.0032)$ 
                }
                \\

                % \input{../../tables/tva_replication_slides.tex}
                \bottomrule
            \end{tabular}

@@

\section{Conclusion}

\begin{frame}{To-do:}
    \begin{itemize}
        \item Include Monte-Carlo simulations on misspecification of spillovers

        \item Extend bias to settings with differential timing of treatment
    \end{itemize}
\end{frame}

\begin{frame}
    \begin{center}
        \Large \textbf{Thank you!}
    \end{center}
\end{frame}
